---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# gguke

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

{gguke} will provide tools for visualizing ukelele chords with lyrics. This is for people that aren't so good at ukelele; they may not really know the chords yet.  It's a chord-first approach to song charts, which might have only chord names or small finger position.  

And because we make it with ggplot2, we can do a lot with customization.  For example, fingers get assigned a color using the viridis palette. 

This is currently a [{readme2pkg}]() project; we're writing in a kind-of stream of consciousness way, bundling up useful code into functions as we go in our README.rmd narrative, and then sending written-up functions to our R folder with knitr/rmd magic. Our commitment to keeping things as they are is low at this point, but it is a real package.  Lifecycle is *early* experimental. See also [{litr}](). 

Beyond the descriptions of our work, we interject comments on our hesitations 🤔 and areas that need some work 🚧, for your consideration marked with emoji.

### First, something I'm very excited about, a new-chord ingestion proposal... - it's just text!

```{r cars}
CM <- 
"pppp
----
----
---3"


library(tidyverse)
```


### Then translate that to data frame that records string, fret, and finger that will play each note.

```{r}
parse_chord <- function(chord = CM){

play_TF <- chord |> stringr::str_split("") %>%  .[[1]] %>% .[1:4]

chart <- chord |> stringr::str_split("") %>%  .[[1]] %>% .[5:length(.)] 
num_frets <- length(chart)/5
string <- rep(1:4, num_frets)
fret <- sort(rep(1:num_frets, 4))

data.frame(finger = chart) %>% 
  dplyr::filter(finger != "\n") %>% 
  dplyr::mutate(fret = fret) %>% 
  dplyr::mutate(string = string) %>% 
  dplyr::mutate(finger = ifelse(.data$finger == "-", NA, finger) %>% as.numeric()) %>% 
  dplyr::filter(!is.na(finger)) ->
fingering_df

list(play_tf = play_TF, 
     fingering_df = fingering_df)

}
```

```{r}

# play or not for each string
parse_chord()[[1]]

# dataframe with finger placement info
parse_chord()[[2]]


```

🤔 maybe read.delim(delim = ""), would work better and feel more grokable that this string split business! :-)



```{r, eval = T}
usethis::use_pipe()
readme2pkg::chunk_to_r("parse_chord")
```

# then we make a fret board. 

```{r}
library(ggplot2)
ggplot() + 
  annotate(geom = "segment", x = 1:4, y = .5, 
           xend = 1:4, yend = 5, linewidth = 3) + 
  annotate(geom = "segment", y = 0:4 +.5, 
           yend = 0:4 +.5, x = 1, xend = 4, linewidth = 3) + 
  scale_y_reverse() + 
  scale_x_continuous(expand = expansion(.2)) +
  theme_void() + 
  coord_equal() +
  scale_fill_viridis_c(limits = c(1,4), guide = F)
```


```{r uke_fretboard}
uke_fretboard <- function(){
  
  ggplot() + 
  annotate(geom = "segment", x = 1:4, y = .5, 
           xend = 1:4, yend = 5, linewidth = 3) + 
  annotate(geom = "segment", y = 0:4 +.5, 
           yend = 0:4 +.5, x = 1, xend = 4, linewidth = 3) + 
  scale_y_reverse() + 
  scale_x_continuous(expand = expansion(.2)) +
  theme_void() + 
  coord_equal() +
  scale_fill_viridis_c(limits = c(1,4), guide = F) 
  
}
```

🤔 maybe gguke() would be better
🤔 maybe a coord_uke could be a long-term goal.

```{r, eval = T}
readme2pkg::chunk_to_r("uke_fretboard")
```


## now we'll place fingers positions w/ point and text

```{r}
uke_fretboard() + 
  geom_point(data = parse_chord()[[2]],
             size = 15, 
             aes(x = string, y = fret),
             color = "white"
             ) +
  geom_point(data = parse_chord()[[2]],
             size = 15, pch = 21, alpha = .6,
             aes(x = string, y = fret,
                 fill = finger), 
             ) +
  geom_text(data = parse_chord()[[2]],
             size = 10,
             aes(x = string, y = fret, label = finger)
             )
```


### looks great! Make a function 

```{r add_chord_fingering}
add_chord_fingering <- function(chord = CM){
  
   list(  geom_point(data = parse_chord(chord)[[2]],
             size = 15,
             aes(x = string, y = fret),
             color = "white"
             ) ,
  geom_point(data = parse_chord(chord)[[2]],
             size = 15, pch = 21, alpha = .6,
             aes(x = string, y = fret,
                 fill = finger),
             ) ,
  geom_text(data = parse_chord(chord)[[2]],
             size = 10,
             aes(x = string, y = fret, label = finger)
             )
   )
}
```  

```{r, eval = T}
readme2pkg::chunk_to_r("add_chord_fingering")
```

## Okay, we want to pair lyrics with the chord

```{r} 
uke_fretboard() +
  add_chord_fingering() +
  labs(caption = "Come stop your cryin', it'll be all right" %>% str_wrap(28)) + 
  theme(text = element_text(size = 23))
```


### looks good... make it a function! 

```{r add_lyric}

add_lyric <- function(lyric = "Come stop your cryin', it'll be all right"){
  
    list(labs(caption = lyric %>% str_wrap(28)),
         theme(text = element_text(size = 23)))
  
  
}
```


```{r, eval = T}
readme2pkg::chunk_to_r("add_lyric")
```

### try it out...

```{r}
uke_fretboard() + 
  add_chord_fingering(CM) + 
  add_lyric("Just take my hand and hold it tight")
```  

### and with new chord input

```{r}

FM <- 
"pppp
--1-
2---
----"

uke_fretboard() + 
  add_chord_fingering(FM) + 
  add_lyric("I will protect you from all around you")
```


# and a bunch more chords and lyrics!

```{r}
Dm <- 
"pppp
--1-
23--
----"


uke_fretboard() + 
  add_chord_fingering(Dm) + 
  add_lyric("I will be here don't you ...")


GM <- 
"pppp
----
-1-2
--3-"



E7M <- 
"pppp
1---
-2-3
----"


AM <- 
"pppp
-1--
2---
----"



DM <- 
"pppp
----
123-
----"

Dbm <- 
"pppp
12--
----
----"

Fsm <- 
"pppp
-1--
2-3-
----"

```


## Lets include the chords as data in the package.  We can add some more later.  Also, you can overwrite what's here, if there is a better fingering. 

Would like to use # but using s for sharp since # is a special character... Feels a little wierd.

```{r}
usethis::use_data(CM, overwrite = T)
usethis::use_data(FM, overwrite = T)
usethis::use_data(Dm, overwrite = T)
usethis::use_data(GM, overwrite = T)
usethis::use_data(E7M, overwrite = T)
usethis::use_data(AM, overwrite = T)
usethis::use_data(DM, overwrite = T)
usethis::use_data(Dbm, overwrite = T)
usethis::use_data(Fsm, overwrite = T)
```


```{r data}
"CM"
"FM"
"Dm"
"GM"
"E7M"
"AM"
"DM"
"Dbm"
"Fsm"
```

# next steps...

```{r}
library(tidyverse)
library(gguke)
lyric_chord_df <- tibble::tribble(~lyric, ~chord_name,
        "Come stop your cryin', it'll be all right", "CM",
"Just take my hand, hold it tight", "CM",
"I will protect you, from all around you", "FM",
"I will be here, don't you ","Dm",
"...cry" , "GM" ,
" ", "E7M",
"'Cause you'll be in my ... heart", "AM",
"... heart", "DM",
"Yes, you'll be in my heart", "E7M",
"... heart", "Dbm", 
"From this day on", "AM",
"Now and forever... more", "DM",
"... more", "GM",
"...", "E7M",
"You'll be in my ...", "AM",
"... heart", "DM",
"No matter what they ...", "E7M",
"say", "Dbm", 
"You'll be here in my... ", "AM", 
" heart", "DM",
"Always", "GM")

lyric_chord_df %>% pull(chord_name) %>% unique()


parse_chord2 <- function(chord){parse_chord(chord)[[2]]}

Dm
lyric_chord_df[1:6,] %>% #just vis first six lyric-chord pairs
  mutate(phrase = row_number()) %>% 
  mutate(fingering_str = map(chord_name, get)) ->
lyric_chord_df_flat


lyric_chord_df_flat %>% 
  unnest(cols = c(fingering_str)) %>% 
  mutate(fingering = map(fingering_str, parse_chord2)) %>% 
  unnest(cols = c(fingering)) ->
lyric_chord_df_fingering
  
lyric_chord_df_fingering %>% 
  ggplot() +
  annotate(geom = "segment", x = 1:4, y = .5, 
           xend = 1:4, yend = 5, linewidth = 1) + 
  annotate(geom = "segment", y = 0:4 +.5, 
           yend = 0:4 +.5, x = 1, xend = 4, linewidth = 1) +
  aes(x = string, y = fret, label = finger) + 
  geom_point(size = 6, color = "white") +
  geom_point(size = 6, pch= 21, aes(fill = finger), alpha = .7 , show.legend = F) +
  geom_text() +
  facet_wrap(~ paste0("phrase ", phrase,": ", chord_name) %>% fct_inorder()) + 
  geom_text(data = lyric_chord_df_flat,
            x = 1, y = -5.5, size = 3,
            aes(label = lyric %>% str_wrap(28)),
            hjust = 0,
            vjust = 1) + 
  scale_y_reverse(limits = c(7, .5)) + 
  coord_equal() + 
  scale_fill_viridis_c(end = .9) + 
  scale_x_continuous(expand = expansion(.6)) + 
  theme_void()

ggwipe::last_plot_wipe(index = 4) + 
    geom_point(size = 6, pch= 21, fill = alpha("white", 0), show.legend = F) 


```

### excited to put in a function!  

```{r lyric_chord_df_to_chart}

parse_chord2 <- function(chord){parse_chord(chord)[[2]]}


lyric_chord_df_to_chart <- function(lyric_chord_df, bw = F){
  
  lyric_chord_df %>% 
  mutate(phrase = row_number()) %>% 
  mutate(fingering_str = map(chord_name, get)) ->
lyric_chord_df_flat

lyric_chord_df_flat %>% 
  unnest(cols = c(fingering_str)) %>% 
  mutate(fingering = map(fingering_str, parse_chord2)) %>% 
  unnest(cols = c(fingering)) ->
lyric_chord_df_fingering
  
lyric_chord_df_fingering %>% 
  ggplot() +
  annotate(geom = "segment", x = 1:4, y = .5, 
           xend = 1:4, yend = 5, linewidth = 1) + 
  annotate(geom = "segment", y = 0:4 +.5, 
           yend = 0:4 +.5, x = 1, xend = 4, linewidth = 1) +
  aes(x = string, y = fret, label = finger) + 
  geom_point(size = 6, color = "white") +
  geom_point(size = 6, pch= 21, aes(fill = finger), alpha = .7 , show.legend = F) +
  geom_text() +
  facet_wrap(~ paste0("phrase ", phrase,": ", chord_name) %>% fct_inorder()) + 
  geom_text(data = lyric_chord_df_flat,
            x = 1, y = -5.5, size = 3,
            aes(label = lyric %>% str_wrap(28)),
            hjust = 0,
            vjust = 1) + 
  scale_y_reverse(limits = c(7, .5)) + 
  coord_equal() + 
  scale_fill_viridis_c(end = .9) + 
  scale_x_continuous(expand = expansion(.6)) + 
  theme_void()


if(bw){
ggwipe::last_plot_wipe(index = 4) + 
    geom_point(size = 6, pch= 21, fill = alpha("white", 0), show.legend = F) 
}
  

}
```

```{r}
readme2pkg::chunk_to_r("lyric_chord_df_to_chart")
```


```{r}
lyric_chord_df_to_chart(lyric_chord_df[1:6,], bw = T)
lyric_chord_df_to_chart(lyric_chord_df[7:12,], bw = T)
lyric_chord_df_to_chart(lyric_chord_df[13:18,], bw = T)
lyric_chord_df_to_chart(lyric_chord_df[19:21,], bw = T)

a_cool_song <- tibble::tribble(~lyric, ~chord_name,
                       "This is my cool lyric", "CM",
                       "This is another", "DM",
                       "hi", "AM")

lyric_chord_df_to_chart(lyric_chord_df = a_cool_song)
```


# Part II. Packaging and documentation  🚧 ✅ 

## Phase 1. Minimal working package

### Created files for package archetecture with `devtools::create(".")` ✅ 

### Moved functions R folder? ✅  

```{r}
knitr::knit_code$get() |> names()
```

Use new {readme2pkg} function to do this from readme if you haven't already

```{r, eval = F}
readme2pkg::chunk_to_r("")
```

### Added roxygen skeleton? ✅ 

Use a roxygen skeleton for auto documentation and making sure proposed functions are *exported*.

### Managed dependencies ? ✅ 

Package dependencies managed, i.e. `depend::function()` in proposed functions and declared in the DESCRIPTION

```{r}
usethis::use_package("ggplot2")
usethis::use_package("dplyr")
```



### Chosen a license? ✅ 


```{r, eval=F}
usethis::use_mit_license()
```

### Run `devtools::check()` and addressed errors? ✅ 

```{r, results='hide', error=T}
devtools::check(pkg = ".")
```

### Build package 🚧 

```{r}
devtools::build()
```

You need to do this before Part 0 in this document will work.

### Make aspirational part of readme real. 🚧 

At this point, you could change eval chunk options to TRUE. You can remove  the 🦄 emoji and perhaps replace it with construction site if you are still uncertain of the API, and want to highlight that it is subject to change. 

### Add lifecycle badge (experimental)

```{r}
usethis::use_lifecycle_badge("experimental")
```


## Phase 2: Listen & iterate 🚧 

Try to get feedback from experts on API, implementation, default decisions.  Is there already work that solves this problem? 


## Phase 3: Let things settle

### Settle on examples.  Put them in the roxygen skeleton and readme. 🚧 

### Written formal tests of functions? 🚧 


That would look like this...

```{r test_calc_frequency_works, eval=F}
library(testthat)

test_that("calc frequency works", {
  expect_equal(calc_frequency("A", 0), 440)
  expect_equal(calc_frequency("A", -1), 220)
  
})
```


```{r, eval = F}
readme2pkg::chunk_to_tests_testthat("test_calc_frequency_works")
```


### Have you worked added a description and author information in the DESCRIPTION file? 🚧 

### Addressed *all* notes, warnings and errors. 🚧 

## Promote to wider audience...

### Package website built? 🚧 

### Package website deployed? 🚧 

## Phase 3: Harden/commit

### Submit to CRAN? Or don't.  🚧 

# Appendix: Reports, Environment

## Description file extract

```{r}

```

## Environment

Here I just want to print the packages and the versions

```{r}
all <- sessionInfo() |> print() |> capture.output()
all[11:17]
```

## `devtools::check()` report

```{r, error = T, results="hide", warning=F}
devtools::check(pkg = ".")
```

```{r}
devtools::build(".")
```

# test functions from package, use :: (or :::)

```{r}
library(gguke)
Fsm

gguke:::uke_fretboard() + 
  gguke:::add_chord_fingering() + 
  gguke:::add_lyric()


gguke:::uke_fretboard() + 
  gguke:::add_chord_fingering(Fsm) + 
  gguke:::add_lyric("a different lyric")




gguke:::lyric_chord_df_to_chart(lyric_chord_df)

lyric_chord_df_to_chart(lyric_chord_df)

```



